The \dealii Library, Version 9.7

### 2025-08-16 9.7/paper.tex
Make grammar unambiguous in one place :-)
tbb vs taskflow results

A large number of operations within \dealii{} are parallelized,
including through mechanisms like WorkStream \cite{TKB16} that are
also available to and useful in application codes. For more than
15 years, since the 6.3 release, \dealii{} has used the Threading
Building Blocks (TBB) \cite{reinders2007tbb} as the underlying library to
provide task- and pipeline-based parallelism. However, the TBB has
downsides that primarily follow from its use of assembler mnemonics
and the resulting difficulties with supporting all platforms and
compilers. As a consequence, starting with the current release,
\dealii{} builds on the Taskflow library \cite{huang2021taskflow} that
is entirely implemented in C++, is header-only, and provides an
interface that matches well with current C++ language standards. \dealii{}
bundles the Taskflow 3.10 release, and now no longer bundles the TBB,
though the TBB interfaces are still available if a TBB installation
is found on a system.

\todo[inline]{Timo: Perhaps add the figures from the WorkStream paper
  updated for TBB vs Taskflow.}

@@

The 9.7 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}
     {around 25 incompatible changes};
see \cite{changes97}. Many of these
incompatibilities remove previously deprecated functions or classes,
or require now widely available but newer versions of external
dependencies than previous \dealii{} versions. Others
change internal
interfaces that are not usually used in external
applications. The following are worth mentioning since they
are more broadly visible:
\begin{itemize}
  \item The classes \texttt{Subscriptor} and \texttt{SmartPointer} have been
    renamed to \texttt{EnableObserverPointer} and \texttt{ObserverPointer}. Many
    core classes, such as \texttt{DoFHandler} and \texttt{Triangulation},
    inherit from \texttt{EnableObserverPointer} to enable other classes to store
    \texttt{ObserverPointer}s pointing to them which, if dereferenced after the
    pointed-to object goes out of scope, will signal an error that the
    pointer is now ``dangling''. This
    functionality has been in \dealii since 1998. Since ``smart pointers''
    nowadays typically refers to the \texttt{std::unique\_ptr} or
    \texttt{std::shared\_ptr} classes, these
    names were updated to avoid this conflict in terminology. The new names are
    based on the proposal for \texttt{std::observer\_ptr}~\cite{Brown2014} and
    \texttt{std::enable\_shared\_from\_this}.

>>>>

A large number of operations within \dealii{} are parallelized,
including through mechanisms like WorkStream \cite{TKB16} that are
also available to and useful in application codes. For more than
15 years, since the 6.3 release, \dealii{} has used the Intel Threading
Building Blocks (TBB) \cite{reinders2007tbb} as the underlying library to
provide task- and pipeline-based parallelism. However, the TBB has
downsides that primarily follow from its use of assembler mnemonics
and the resulting difficulties with supporting all platforms and
compilers. As a consequence, starting with the current release,
\dealii{} builds on the Taskflow library \cite{huang2021taskflow} that
is entirely implemented in C++, is header-only, and provides an
interface that matches well with current C++ language standards. \dealii{}
bundles the Taskflow 3.10 release, and now no longer bundles the TBB,
though the TBB interfaces are still available if a TBB installation
is found on a system.

In Figure~\ref{fig:tbb-vs-taskflow} we compare the time for assembling
the temperature matrix and temperature right-hand side in the step-32
tutorial program, like it was done in \cite{TKB16}. The test was performed
on a 3d mesh with 393,216 cells and 3,195,010 temperature degrees of freedom
using a $Q_2$ element. The tests are run
on AMD EPYC 9654P with 96 cores (and 192 threads). One can clearly see that
running all copy operations sequentially limits parallel scalability (to less than 10
threads on the left, where the copy operation is more expensive than the right),
while using graph coloring (denoted as ``colored'') eliminates this bottleneck.
It is clearly visible that the Taskflow implementation is better or the same
in all situations.

\begin{figure}[hb]
\label{fig:tbb-vs-taskflow}
\includegraphics{kokkos-vs-tbb/plots-figure1.pdf}\hfill
\includegraphics{kokkos-vs-tbb/plots-figure2.pdf}
\caption{Assembly times in a modified step-32 (left: temperature matrix assembly, right: temperature right-hand side assembly) comparing TBB and Taskflow with and without coloring.}
\end{figure}

@@

The 9.7 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}
     {around 25 incompatible changes};
see \cite{changes97}. Many of these
incompatibilities remove previously deprecated functions or classes,
or require now widely available but newer versions of external
dependencies than previous \dealii{} versions. Others
change internal
interfaces that are not usually used in external
applications. The following are worth mentioning since they
are more broadly visible:
\begin{itemize}
  \item The classes \texttt{Subscriptor} and \texttt{SmartPointer} have been
    renamed to \texttt{EnableObserverPointer} and \texttt{ObserverPointer}. Many
    core classes, such as \texttt{DoFHandler} and \texttt{Triangulation},
    inherit from \texttt{EnableObserverPointer} to enable other classes to store
    \texttt{ObserverPointer}s pointing to them which, if dereferenced after the
    pointed-to object goes out of scope, will signal an error that the
    pointer is now ``dangling''. This
    functionality has been in \dealii since 1998. Since the term ``smart pointers''
    nowadays typically refers to the \texttt{std::unique\_ptr} or
    \texttt{std::shared\_ptr} classes, these
    names were updated to avoid this conflict in terminology. The new names are
    based on the proposal for \texttt{std::observer\_ptr}~\cite{Brown2014} and
    \texttt{std::enable\_shared\_from\_this}.
